JD notes on effects

### 2022-07-11 manuscript/body.tex
More progress, I guess.

\jd{It looks like you're trying to introduce column averaging and non-linear averaging in the same \P. How can we simplify?}
For linear models (including models with complex interactions), the averaging is done on the linear scale, i.e., linear averaging. As a result, the \emph{bias} in the expected (mean) predictions can easily be corrected using \emph{mean-based reference point}, i.e., averaging of the non-focal model variables -- currently not implemented in commonly used \proglang{R} software packages. The mean of the central estimates is the prediction at the \emph{model center}. In this case, the estimates are usually consistent with the observed values. However, in models with non-linear link functions, this is not always the case. In particular, when dealing with GL(M)Ms with non-linear link functions, the averaging is done on the non-linear scale, i.e., non-linear averaging, making generating correct predictions much harder. This is the bias in the expected mean prediction induced by the non-linear transformation of the response variable. One way to address this is to make predictions on the linear predictor scale and back-transform to the original scale. However, the back-transformation may either result in biased predictions or requires some approximation, for example, second-order Taylor approximation implemented in \pkg{emmeans} \citep{lenth2018package}. An alternative to the mean-based reference point is the \emph{whole-sample-based} approach, discussed in detail later, which involves computing the prediction over the population of non-focal predictors and then averaging across the values of the focal predictor \citep{hanmer2013behind}.

This article aims to discuss and implement various approaches for computing predictions and effects, together with the associated plots. We further explore and demonstrate, using simulated data, approaches for correcting bias in central estimates for GL(M)Ms involving non-linear link functions. The proposed method and \proglang{R} software package will complement the existing ones by providing: 1) a straightforward way to generate effects, and 2) an alternative and a more robust way to correct for prediction bias in GL(M)Ms.

@@

\item \textbf{Model variables}: Refer to the variables which are combined to make the linear predictor (corresponding to the columns in the model matrix). Each input variable may correspond to one or more model variable. Input variables with more than two categories (for instance religion in our previous example), or input variables with non-linear response functions (e.g., spline or polynomial) will correspond to more than one model variable -- we call such variables “multi-parameter variables”.

\item \textbf{Model center:} A point corresponding to a column-wise mean of the model matrix (the mean of one or more model variables). Also referred to as center point. The center point for a set of model variables corresponding to an input variable may not represent a possible value of the input variable -- the case in \FREF{justify_plots}B. In simple linear models with no interactions, the mean of the input variable can be used as a proxy for the center point \jd{Is it a “proxy” or is it the thing??}. If there are more than one model variables associated with the focal input variable, however, the model center does not correspond to any given value of the focal predictor \jd{This seems like bad repetition}, and we do not necessarily get the ``nice'' crossing we get in simpler cases, see \FREF{pred_cubic_plots}.

@@

\item \textbf{Anchor:} The value chosen for the focal predictor when estimating effect-style confidence intervals. The anchor choice does not affect the central estimates, nor prediction-style. Often chosen as the center point of the model variables corresponding to the focal predictor.

@@

Multi-parameter variables (MPVs) such as splines, polynomials, etc., can be within the focal predictor, or within non-focal predictor(s). To distinguish the two, suppose the model which describes the hypothetical simulation of household size based on a number of socio-demographic factors such as age and wealth index is
%
\begin{align}\label{eq:lm_cubic}
\mathrm{hh~size}_i &= \beta_0 + \beta_{\mathrm{A_1}}\mathrm{Age}_i + \beta_{\mathrm{A_2}}\mathrm{Age}^2_i + \beta_{\mathrm{A_3}}\mathrm{Age}^3_i\nonumber \\
&+ \beta_{\mathrm{W}}\mathrm{Wealthindex}_i + \epsilon_i.
\end{align}
%
In the first case, with \code{Age} as the focal predictor, is a cubic polynomial with three model variables ($\mathrm{Age}_i, \mathrm{Age}^2_i$ and $\mathrm{Age}^3_i$). In this case, each model variable is evaluated separately across the chosen levels of the focal predictor, $Age_i$. Specifically, the model variables are treated as additional columns of the model matrix evaluated with the same values chosen for the focal predictor, while non-focal predictors are fixed at their reference point as discussed in the previous section. The central estimates associated with \code{Age} on the linear predictor scale become
%
\begin{align}
\eta^\star(\mathrm{Age}_i, \nset{{\over{\mathrm{Wealthindex}}}}) &= \beta_0 + \beta_{\mathrm{A_1}}\mathrm{Age}_i + \beta_{\mathrm{A_2}}\mathrm{Age}^2_i + \beta_{\mathrm{A_3}}\mathrm{Age}^3_i \nonumber\\
	& + \beta_{\mathrm{W}}\over{\mathrm{Wealthindex}}.
\end{align}
%
In the second case, with \code{Wealthindex} as the focal predictor, the non-focal predictor, \code{Age}, is a cubic polynomial. In this case, the non-focal model variables ($\mathrm{Age}_i, \mathrm{Age}^2_i$ and $\mathrm{Age}^3_i$) are all treated as separate non-focal model variables and an appropriate choice of reference point applies just like in the models without MPVs. For instance, in mean-based approach, we average all non-focal MPVs. Thus
%
\begin{align}
\eta^\star(\mathrm{Wealthindex}_i, \nset{{\{\over{\mathrm{Age\mathop{\vphantom{^2}}}}, \over{\mathrm{Age}^2}, \over{\mathrm{Age}^2}\}}}) &= \beta_0 + \beta_{\mathrm{A_1}}\over{\mathrm{Age\mathop{\vphantom{^2}}}} + \beta_{\mathrm{A_2}}\over{\mathrm{Age}^2} + \beta_{\mathrm{A_3}}\over{\mathrm{Age}^3}\nonumber\\
	& + \beta_{\mathrm{W}}\mathrm{Wealthindex}_i.
\end{align}
%

We could also have interactions between focal and non-focal predictors. To illustrate, we consider simple input variables but the logic extends to more complex ones like MPVs. Consider model described by \EREF{simple_inter_higher}, used to demonstrate how to handle interactions in non-focal predictors. Here, we use it to demonstrate how to handle interactions between the focal, $x_2$, and non-focal, $x_3$, predictors. In this case, the non-interacting predictors are treated as non-focal model variables, with appropriate choice of reference points. However, for the interacting model variables, we first choose representative values for the focal predictor and then some particular values of the interacting non-focal predictors in a similar way we choose the values of the focal predictor or use predetermined values. In our example, suppose we pick $i$ and $j$ unique values of the focal predictor, $x_2$ and interacting non-focal predictor, $x_3$, respectively. The central estimate on linear predictor scale is given by
%
\begin{align*}
\eta^\star(x_{2i}, x_{3j}, {\bar{x}^\star_1}) = \beta_0 + \beta_1 \bar{x}^\star_1 + \beta_2x_{2i} + \beta_3x_{3j} + \beta_{23}x_{2i}x_{3j}.
\end{align*}
%

@@

We describe the uncertainty around the estimates using confidence intervals (CIs). In principle, every value of focal predictor has a different CI. The conventional way to compute variances for predictions is 
%
\begin{align}\label{eq:conventional_variance}
\boldsymbol\sigma^2 = \textrm{Diag}(\bX^\star \boldsymbol{\Sigma} \bX^{\star\top}), 
\end{align}
so that the confidence intervals are $\boldeta \pm q\boldsymbol\sigma$, where $q$ is an appropriate quantile of Normal or t distribution \citep{lenth2018package, fox2009effect}. This generates conventional CIs which incorporate all the uncertainties -- including the uncertainties due to the intercept and non-focal predictor.  But what if we are interested in the uncertainty as a result of the focal predictor only, so that the CIs are $\boldeta \pm q \boldsymbol \sigma_f$, i.e., effects?

\NEW{We may be interested in the uncertainties associated with the focal predictor only, excluding other uncertainties due to other non-focal predictors -- effects. However, commonly used \proglang{R} packages for constructing predictions do not exclude the uncertainties resulting from non focal predictors when computing the CIs.} Currently obscure way to exclude uncertainties associated with non-focal predictors in some of these packages is to provide a user defined variance-covariance matrix with the covariances of non-focal predictors set to $0$. We refer to this procedure as \emph{zeroing-out} variance-covariance matrix; and only works when the input variables are \emph{centered} prior to model fitting, in case of numerical variables, and more complicated when the input variables are categorical. An alternative is our proposed approach which uses the model center and does not require the input variables to be centered prior to model fitting.

Let $\bxo$ be a centered model matrix previously defined, and let $\ba$ be an anchor matrix, with the same dimensions and entries in all non-focal predictors as $\bxo$. Let $\baf$ be the column of $\ba$ corresponding to focal predictor(s) defined in $\bxo$. Any appropriate values can be chosen for $\baf$ but for model center (center-anchored), we use $\bafc$ which is the mean of the focal predictor(s). Thus 
%
\begin{align}\label{eq:centered_variance}
\boldsymbol\sigma_f^2 = \textrm{Diag}((\bxo - \bafc) \boldsymbol{\Sigma} (\bxo - \bafc)^\top).
\end{align}
%
We can see that $\forall ~\bxof=\baf$, $\bxo - \baf = \boldsymbol 0$, hence $\boldsymbol\sigma_f^2 = \boldsymbol{0}$. Similarly, for all values of $\bxof$ close to anchor point $\bafc$, the term $(\bxo - \bafc)$ and $\boldsymbol\sigma_f^2$ goes to $\boldsymbol 0$ and $\boldsymbol\sigma_f^2 = \boldsymbol 0$ if $\bxof=\bafc$. This means that $\boldsymbol\sigma_f^2$ close to the anchor point are smaller than those away from the anchor point; and results to confidence intervals which are narrower around the anchor or crosses at the anchor point for simple models. This is the effect of the particular focal predictor. By setting $\baf = \boldsymbol 0$, we get the variances for the conventional CIs in \EREF{conventional_variance}.

The computation of $\bxo - \bafc$ impacts only on the intercepts and non-focal predictors, i.e., the slopes and variance corresponding to the focal predictors are not affected. This means that we can still generate effects without necessarily centering the predictors prior to model fitting.

@@

We can directly compare the outcome estimate at the model center (mean-based estimate) with the average of predictions evaluated across the population values of non-focal predictors (sample-based estimate):

@@

In an ordinary linear model, the link function $g$ is the identity function, so the two means are the same. For non-trivial link functions, we expect them to be different in general -- \NEW{a consequent of Jensen inequality \citep{duursma_bias_2003, denny_performance_2019, ye_generalized_2021}. In GL(M)Ms the assumed link function, predetermines the curvature of the response surface and also the sign of the Jensen effect. For example, logistic model has a positive curvature if the predicted probabilities are low and negative curvature if the predicted probabilities are high. In other words, logistic model may be concave or convex depending on the predicted probabilities, hence the sign of the Jensen effect is not readily available. However, in log link functions, the inverse of the link (exponential) function is strictly convex, i.e., an accelerating function. For accelerating functions, the Jensen’s inequality describes how the changes in input values elevate the predictions and otherwise describes how these changes depress the predictions. Further, our models may include input variables and/or random effects terms other than the focal one. In non-trivial link functions, the Jensen effect may depend on the additional input variables \citep{ye_generalized_2021}}.

\NEW{It is usually important to report the estimates that reflect the expected values of the untransformed response. This is not a major issue in simple linear models. However, when dealing with nonlinear link functions,  generated predictions may not reflect the observed response due to the bias in the expected mean induced by the nonlinear transformation of the response variable -- a consequent of Jensen effect mentioned above. In such cases, bias correction is needed to back-transform the predictions to the original scales. The common approach for bias-adjustment is second-order Taylor approximation \citep{duursma2003bias, hanmer2013behind}; already implemented in \pkg{emmeans} \citep{lenth2018package}. Here, we describe and implement a different approach -- whole-sample-based approach for bias correction. Hammer and Kalkan \citep{hanmer2013behind} introduced mean- and whole-sample- based approaches but limited their discussion to binary response models only.}

@@

The most commonly used \proglang{R} software packages for outcome plots (\pkg{emmeans} and \pkg{effects}), by default, use the average of input variables and ``sum-to-zero'' as the reference point approach for continuous and categorical input variables, respectively. However, there are a number of choices one can make when constructing outcome plots -- for example, in the presence of interactions, averaging the interactions (averaging the product of model variables) versus taking the product of the averages of the input variables (the default for \pkg{emmeans} and \pkg{effects}). For simple OLS models (with no interactions), the two approaches give the same estimates. However, for complex models, we demonstrate that the two approaches can produce substantially different results and show that averaging the interactions closely matches the observed values. To illustrate this, consider models~\ref{eq:simple_inter_higher_no_interaction} and \ref{eq:simple_inter_higher} below, with $x_1$ as the focal predictor:

@@

In the absence of interactions (\EREF{simple_inter_higher_no_interaction}), the three approaches produce the same estimates, which match the simulated values, \FREF{justify_plots}A. However, in the presence of interactions, even as simple as the one in \EREF{simple_inter_higher}, the estimates start to differ. In particular, \pkg{emmeans} and \pkg{effects} have similar estimates ($\bar{\hat{y}}$) but different from the \pkg{varpred}'s which, however, is very close to the simulated average ($\bar{y}$), \FREF{justify_plots}B. To generate \FREF{justify_plots}A,  \pkg{emmeans} and \pkg{effects} average $x_2$ and $x_3$, i.e., the input variables while \pkg{varpred} averages the model variables corresponding to $x_2$ and $x_3$. On other hand, to generate \FREF{justify_plots}B, the difference in the estimates lies on how each of the packages average the interaction model variable ($x_2x_3$). In particular, \pkg{emmeans} and \pkg{effects} use input variables to compute $\bar{x_2}\bar{x_3}$ while \pkg{varpred} uses model variables to compute $\over{x_2x_3}$. In the case of \FREF{justify_plots}A, averaging the input variables is equivalent to averaging the model variables.

@@

The \proglang{R} software package \pkg{margins} \citep{lenth2018package} generates marginal effects. To illustrate the distinction between effects and marginal effects, we use the simulated data described in \nameref{S1_Appendix}. 
%
\begin{figure}
\centering
\includegraphics{cubic_varpred_margins_pred.ggp.pdf}
\caption{{\bf Effect and marginal effect plots.} \NEW{A:  The effect plot. B: The marginal effect (ME) at representative values of the focal predictor \code{age} together with the $95\%$ confidence intervals. For simple linear models with no interactions, the ME should be constant across all the values of the focal predictor; and it would correspond to the slope of the central estimate of central estimate. However, in this example, the focal predictor is a cubic polynomial, and so, the effect is not constant but a function of the partial derivatives of the corresponding model variables.}}
\label{fig:qoi_age_pred_plot}
\end{figure}
%

@@

\begin{figure}
\begin{center}
\includegraphics{cubic_predictors_preds.ggp.pdf}
\end{center}
\caption{{\bf Prediction and effect plots for cubic polynomial and simple focal predictors.} A: The focal predictor is a MPV cubic polynomial. B: The focal input variable is linear, with MPV non-focal predictor. The central estimates (dashed central curves or lines) are the same for predictions and effects, in each case (A and B). However, the effects are narrower than the predictions. The default anchor for generating effects is the center point. For linear (or simple) focal input variables, the center point is a point within the range of focal values; and therefore the effect curves intersect at that that point -- B. In the case of MPVs such as cubic polynomials, the center point is not a point but a combination of all the model variables associated with the cubic polynomial, consequently, curves will not necessarily cross at a point -- A. The horizontal black and red dashed lines are the observed and predicted average household size, i.e., $\over{hh~size}$ and $\over{\widehat{hh~size}}$, respectively.}
\label{fig:pred_cubic_plots}
\end{figure}

>>>>

\jd{It looks like you're trying to introduce column averaging and non-linear averaging in the same \P. How can we simplify?}
For linear models (including models with complex interactions), the averaging is done on the linear scale, i.e., linear averaging. As a result, the \emph{bias} in the expected (mean) predictions can easily be corrected using \emph{mean-based reference point}, i.e., averaging of the non-focal linear predictor variables -- currently not implemented in commonly used \proglang{R} software packages. The mean of the central estimates is the prediction at the \emph{model center}. In this case, the estimates are usually consistent with the observed values. However, in models with non-linear link functions, this is not always the case. In particular, when dealing with GL(M)Ms with non-linear link functions, the averaging is done on the non-linear scale, i.e., non-linear averaging, making generating correct predictions much harder. This is the bias in the expected mean prediction induced by the non-linear transformation of the response variable. One way to address this is to make predictions on the linear predictor scale and back-transform to the original scale. However, the back-transformation may either result in biased predictions or requires some approximation, for example, second-order Taylor approximation implemented in \pkg{emmeans} \citep{lenth2018package}. An alternative to the mean-based reference point is the \emph{whole-sample-based} approach, discussed in detail later, which involves computing the prediction over the population of non-focal predictors and then averaging across the values of the focal predictor \citep{hanmer2013behind}.

This article aims to discuss and implement various approaches for computing predictions and effects, together with the associated plots. We further explore and demonstrate, using simulated data, approaches for correcting bias in central estimates for GL(M)Ms involving non-linear link functions. The proposed method and \proglang{R} software package will complement the existing ones by providing: 1) a straightforward way to generate effects, and 2) an alternative and a more robust way to correct for prediction bias in GL(M)Ms. \jd{We should revisit this list.}

@@

\item \textbf{Linear predictor variables}: Refer to the variables which are combined to make the linear predictor (corresponding to the columns in the model matrix). Each input variable may correspond to one or more model variable. Input variables with more than two categories (for instance religion in our previous example), or input variables with non-linear response functions (e.g., spline or polynomial) will correspond to more than one linear predictor variable -- we call such variables “multi-parameter variables”.

\item \textbf{Model center:} A point corresponding to a column-wise mean of the model matrix (the mean of one or more linear predictor variables). Also referred to as center point. The center point for a set of linear predictor variables corresponding to an input variable may not represent a possible value of the input variable -- the case in \FREF{justify_plots}B. In simple linear models with no interactions, the mean of the input variable can be used as a proxy for the center point \jd{Is it a “proxy” or is it the thing??}. If there are more than one linear predictor variables associated with the focal input variable, however, the model center does not correspond to any given value of the focal predictor \jd{This seems like bad repetition}, and we do not necessarily get the ``nice'' crossing we get in simpler cases, see \FREF{pred_cubic_plots}.

@@

\item \textbf{Anchor:} The value chosen for the focal predictor when estimating effect-style confidence intervals. The anchor choice does not affect the central estimates, nor prediction-style. Often chosen as the center point of the linear predictor variables corresponding to the focal predictor.

@@

\jd{I feel the first s.~is perfectly clear, and I don't know why it needs any explanation at all.}
Multi-parameter variables (MPVs) such as splines, polynomials, etc., can be within the focal predictor, or within non-focal predictor(s). To distinguish the two, suppose the model which describes the hypothetical simulation of household size based on a number of socio-demographic factors such as age and wealth index is
%
\begin{align}\label{eq:lm_cubic}
\mathrm{hh~size}_i &= \beta_0 + \beta_{\mathrm{A_1}}\mathrm{Age}_i + \beta_{\mathrm{A_2}}\mathrm{Age}^2_i + \beta_{\mathrm{A_3}}\mathrm{Age}^3_i\nonumber \\
&+ \beta_{\mathrm{W}}\mathrm{Wealthindex}_i + \epsilon_i.
\end{align}
%
In the first case, with \code{Age} as the focal predictor, is a cubic polynomial with three linear predictor variables ($\mathrm{Age}_i, \mathrm{Age}^2_i$ and $\mathrm{Age}^3_i$). In this case, each linear predictor variable is evaluated separately across the chosen levels of the focal predictor, $Age_i$. Specifically, the linear predictor variables are treated as additional columns of the model matrix evaluated with the same values chosen for the focal predictor, while non-focal predictors are fixed at their reference point as discussed in the previous section. The central estimates associated with \code{Age} on the linear predictor scale become
%
\begin{align}
\eta^\star(\mathrm{Age}_i, \nset{{\over{\mathrm{Wealthindex}}}}) &= \beta_0 + \beta_{\mathrm{A_1}}\mathrm{Age}_i + \beta_{\mathrm{A_2}}\mathrm{Age}^2_i + \beta_{\mathrm{A_3}}\mathrm{Age}^3_i \nonumber\\
	& + \beta_{\mathrm{W}}\over{\mathrm{Wealthindex}}.
\end{align}
%
In the second case, with \code{Wealthindex} as the focal predictor, the non-focal predictor, \code{Age}, is a cubic polynomial. In this case, the non-focal linear predictor variables ($\mathrm{Age}_i, \mathrm{Age}^2_i$ and $\mathrm{Age}^3_i$) are all treated as separate non-focal linear predictor variables and an appropriate choice of reference point applies just like in the models without MPVs. For instance, in mean-based approach, we average all non-focal MPVs. Thus
%
\begin{align}
\eta^\star(\mathrm{Wealthindex}_i, \nset{{\{\over{\mathrm{Age\mathop{\vphantom{^2}}}}, \over{\mathrm{Age}^2}, \over{\mathrm{Age}^2}\}}}) &= \beta_0 + \beta_{\mathrm{A_1}}\over{\mathrm{Age\mathop{\vphantom{^2}}}} + \beta_{\mathrm{A_2}}\over{\mathrm{Age}^2} + \beta_{\mathrm{A_3}}\over{\mathrm{Age}^3}\nonumber\\
	& + \beta_{\mathrm{W}}\mathrm{Wealthindex}_i.
\end{align}
%

\jd{Why is this under MPV as a subhead? Why “also”? Do we discuss non-focal interaction?} We could also have interactions between focal and non-focal predictors. To illustrate, we consider simple input variables but the logic extends to more complex ones like MPVs. Consider model described by \EREF{simple_inter_higher}, \jd{This is referring back to the definition list, which is a strange place to put examples} used to demonstrate how to handle interactions in non-focal predictors. Here, we use it to demonstrate how to handle interactions between the focal, $x_2$, and non-focal, $x_3$, predictors. In this case, the non-interacting predictors are treated as non-focal linear predictor variables, with appropriate choice of reference points \jd{What is the point of saying this again? AFAICT this s.~says if we have a non-interacting predictor we treat it as a non-interacting predictor as already described Can we delete?}. However, for the interacting linear predictor variables, we first choose representative values for the focal predictor and then some particular values of the interacting non-focal predictors in a similar way we choose the values of the focal predictor or use predetermined values. In our example, suppose we pick $i$ and $j$ unique values of the focal predictor, $x_2$ and interacting non-focal predictor, $x_3$, respectively. The central estimate on linear predictor scale is given by
%
\begin{align*}
\eta^\star(x_{2i}, x_{3j}, {\bar{x}^\star_1}) = \beta_0 + \beta_1 \bar{x}^\star_1 + \beta_2x_{2i} + \beta_3x_{3j} + \beta_{23}x_{2i}x_{3j}.
\end{align*}
%
\jd{I don't understand how this part fits in at all. Obviously in theory we could pick values of any non-focal predictor, but what's actually going on here? What's the actual difference between how we pick reference values for interacting vs.~non-interacting non-focal predictors?}

@@

We describe the uncertainty around the estimates using confidence intervals (CIs). In principle, every prediction has a different CI. The conventional way to compute variances for predictions is 
%
\begin{align}\label{eq:conventional_variance}
\boldsymbol\sigma^2 = \mathrm{Diag}(\bX^\star \boldsymbol{\Sigma} \bX^{\star\top}), 
\end{align}
\jd{I don't understand how a scalar ($\sigma^2$) is equal to a vector (the $\mathrm{Diag}$). Is there a citation for this formula?}
so that the confidence intervals are $\boldeta \pm q\boldsymbol\sigma$, where $q$ is an appropriate quantile of Normal or t distribution \citep{lenth2018package, fox2009effect}. This generates conventional CIs which incorporate all the uncertainties -- including the uncertainties due to the intercept and non-focal predictor.  But what if we are interested in the uncertainty as a result of the focal predictor only, so that the CIs are $\boldeta \pm q \boldsymbol \sigma_f$, i.e., effects?

\jd{Haven't we already discussed this idea? The new sentence reads like we didn't. We should start by saying what we want, then continue by explaining that there is an equivalent calculation (is it always equivalent?) that can be done in the packages.}
\NEW{We may be interested in the uncertainties associated with the focal predictor only, excluding other uncertainties due to other non-focal predictors -- effects. However, commonly used \proglang{R} packages for constructing predictions do not exclude the uncertainties resulting from non focal predictors when computing the CIs.} Currently obscure way to exclude uncertainties associated with non-focal predictors in some of these packages is to provide a user defined variance-covariance matrix with the covariances of non-focal predictors set to $0$. We refer to this procedure as \emph{zeroing-out} variance-covariance matrix; and only works when the input variables are \emph{centered} prior to model fitting, in case of numerical variables, and more complicated when the input variables are categorical. An alternative is our proposed approach which uses the model center and does not require the input variables to be centered prior to model fitting.

\jd{You explain it like the 0 variance at the center is the point. The point is to look at the effect of changing the focal parameter from the anchor value; the 0 variance at the anchor value is a corollary. You should start by explaining why something like \EREF{centered_variance} is what you want, and then explain the zero variance.}
\jd{The explanation of \bxo is not clear. I guess you mean that the values for everything except for the focal columns are centered? What is the point of defining \bxo\ and then defining \bafc\ so that it exactly cancels? Is it better to just define a matrix with zeroes in all the right places?}
Let $\bxo$ be a centered model matrix previously defined, and let $\ba$ be an anchor matrix, with the same dimensions and entries in all non-focal predictors as $\bxo$. Let $\baf$ be the column of $\ba$ corresponding to focal predictor(s) defined in $\bxo$. Any appropriate values can be chosen for $\baf$ but for model center (center-anchored), we use $\bafc$ which is the mean of the focal predictor(s). Thus 
%
\begin{align}\label{eq:centered_variance}
\boldsymbol\sigma_f^2 = \textrm{Diag}((\bxo - \bafc) \boldsymbol{\Sigma} (\bxo - \bafc)^\top).
\end{align}
%
We can see that $\forall ~\bxof=\baf$, $\bxo - \baf = \boldsymbol 0$, hence $\boldsymbol\sigma_f^2 = \boldsymbol{0}$. Similarly, for all values of $\bxof$ close to anchor point $\bafc$, the term $(\bxo - \bafc)$ and $\boldsymbol\sigma_f^2$ goes to $\boldsymbol 0$ and $\boldsymbol\sigma_f^2 = \boldsymbol 0$ if $\bxof=\bafc$. This means that $\boldsymbol\sigma_f^2$ close to the anchor point are smaller than those away from the anchor point; and results to confidence intervals which are narrower around the anchor or crosses at the anchor point for simple models. This is the effect of the particular focal predictor. By setting $\baf = \boldsymbol 0$, we get the variances for the conventional CIs in \EREF{conventional_variance}.

The computation of $\bxo - \bafc$ impacts only on the intercepts and non-focal predictors, i.e., the slopes and variance corresponding to the focal predictors are not affected. This means that we can still generate effects without necessarily centering the predictors prior to model fitting. \jd{What is the point of this s.? What does it mean “impacts only on”?}

@@

\jd{Why at the model center? Aren't we interested in comparisons at the reference point for any value of the focal variable?}
We can directly compare the outcome estimate at the model center (mean-based estimate) with the average of predictions evaluated across the population values of non-focal predictors (sample-based estimate):

@@

In an ordinary linear model, the link function $g$ is the identity function, so the two means are the same. For non-trivial link functions, we expect them to be different in general -- \NEW{a consequent of Jensen inequality \citep{duursma_bias_2003, denny_performance_2019, ye_generalized_2021} \jd{Drop this claim; it's obvious we expect them to be different, Jensen is just about the sign}. In GL(M)Ms the assumed link function, predetermines the curvature of the response surface and also the sign of the Jensen effect. For example, logistic model has a positive curvature if the predicted probabilities are low and negative curvature if the predicted probabilities are high. In other words, logistic model may be concave or convex depending on the predicted probabilities, hence the sign of the Jensen effect is not readily available. However, in log link functions, the inverse of the link (exponential) function is strictly convex, i.e., an accelerating function. For accelerating functions, the Jensen’s inequality describes how the changes in input values elevate the predictions and otherwise describes how these changes depress the predictions. Further, our models may include input variables and/or random effects terms other than the focal one. In non-trivial link functions, the Jensen effect may depend on the additional input variables \citep{ye_generalized_2021}}.
\jd{I think I asked you to describe how Jensen is working in these cases. In each of the cases it would be both easier and clearer to say whether we expect the left side or the right side to be larger than to spin all of those words about “describes how the changes in input values elevate the predictions and otherwise describes how these changes depress the predictions”.}

\NEW{It is usually important to report the estimates that reflect the expected values of the untransformed response \jd{Hard to understand this s.}. This is not a major issue in simple linear models \jd{What do you mean by major issue? In what sense is it an issue? In what sense does it exist?}. However, when dealing with nonlinear link functions,  generated predictions may not reflect the observed response due to the bias in the expected mean induced by the nonlinear transformation of the response variable -- a consequent of Jensen effect mentioned above. In such cases, bias correction is needed to back-transform the predictions to the original scales. The common approach for bias-adjustment is second-order Taylor approximation \citep{duursma2003bias, hanmer2013behind}; already implemented in \pkg{emmeans} \citep{lenth2018package}. Here, we describe and implement a different approach -- whole-sample-based approach for bias correction. Hanmer and Kalkan \citep{hanmer2013behind} introduced mean- and whole-sample- based approaches but limited their discussion to binary response models only.} 
\jd{We mostly talk about binary-response models as well. Also, they did not introduce (there are much older citations to it, and they use a different name, which we should mention (or maybe adopt).}

@@

The most commonly used \proglang{R} software packages for outcome plots (\pkg{emmeans} and \pkg{effects}), by default, use the average of input variables and ``sum-to-zero'' as the reference point approach for continuous and categorical input variables, respectively. However, there are a number of choices one can make when constructing outcome plots -- for example, in the presence of interactions, averaging the interactions (averaging the product of linear predictor variables) versus taking the product of the averages of the input variables (the default for \pkg{emmeans} and \pkg{effects}). For simple OLS models (with no interactions), the two approaches give the same estimates. However, for complex models, we demonstrate that the two approaches can produce substantially different results and show that averaging the interactions closely matches the observed values. To illustrate this, consider models~\ref{eq:simple_inter_higher_no_interaction} and \ref{eq:simple_inter_higher} below, with $x_1$ as the focal predictor:

@@

In the absence of interactions (\EREF{simple_inter_higher_no_interaction}), the three approaches produce the same estimates, which match the simulated values, \FREF{justify_plots}A. However, in the presence of interactions, even as simple as the one in \EREF{simple_inter_higher}, the estimates start to differ. In particular, \pkg{emmeans} and \pkg{effects} have similar estimates ($\bar{\hat{y}}$) but different from the \pkg{varpred}'s which, however, is very close to the simulated average ($\bar{y}$), \FREF{justify_plots}B. To generate \FREF{justify_plots}A,  \pkg{emmeans} and \pkg{effects} average $x_2$ and $x_3$, i.e., the input variables while \pkg{varpred} averages the linear predictor variables corresponding to $x_2$ and $x_3$. On other hand, to generate \FREF{justify_plots}B, the difference in the estimates lies on how each of the packages average the interaction linear predictor variable ($x_2x_3$). In particular, \pkg{emmeans} and \pkg{effects} use input variables to compute $\bar{x_2}\bar{x_3}$ while \pkg{varpred} uses linear predictor variables to compute $\over{x_2x_3}$. In the case of \FREF{justify_plots}A, averaging the input variables is equivalent to averaging the linear predictor variables.

@@

The \proglang{R} software package \pkg{margins} \citep{lenth2018package} generates marginal effects. To illustrate the distinction between effects and marginal effects, we use the simulated data described in \nameref{S1_Appendix}. 
%
\begin{figure}
\centering
\includegraphics{cubic_varpred_margins_pred.ggp.pdf}
\caption{{\bf Effect and marginal effect plots.} \NEW{A:  The effect plot. B: The marginal effect (ME) at representative values of the focal predictor \code{age} together with the $95\%$ confidence intervals. For simple linear models with no interactions, the ME should be constant across all the values of the focal predictor; and it would correspond to the slope of the central estimate of central estimate. However, in this example, the focal predictor is a cubic polynomial, and so, the effect is not constant but a function of the partial derivatives of the corresponding linear predictor variables.}}
\label{fig:qoi_age_pred_plot}
\end{figure}
%

@@

\begin{figure}
\begin{center}
\includegraphics{cubic_predictors_preds.ggp.pdf}
\end{center}
\caption{{\bf Prediction and effect plots for cubic polynomial and simple focal predictors.} A: The focal predictor is a MPV cubic polynomial. B: The focal input variable is linear, with MPV non-focal predictor. The central estimates (dashed central curves or lines) are the same for predictions and effects, in each case (A and B). However, the effects are narrower than the predictions. The default anchor for generating effects is the center point. For linear (or simple) focal input variables, the center point is a point within the range of focal values; and therefore the effect curves intersect at that that point -- B. In the case of MPVs such as cubic polynomials, the center point is not a point but a combination of all the linear predictor variables associated with the cubic polynomial, consequently, curves will not necessarily cross at a point -- A. The horizontal black and red dashed lines are the observed and predicted average household size, i.e., $\over{hh~size}$ and $\over{\widehat{hh~size}}$, respectively.}
\label{fig:pred_cubic_plots}
\end{figure}
